---
title: Daphnopsis - Framework
author: "Gabriel Santos"
date: "2024-11-11"
output: html_document
tags:
  - ITV
  - Projetos/ITV/Daphonpsis
aliases:
  - Análises Daphnopsis
  - Script Daphnopsis
---

# Setting

-   Packages
-   Load data

```{r "Packages and loading data", echo=FALSE}
gc()
rm(list=ls())

# Packages
library(openxlsx)
library(tidyverse)

# Data
year2022<-read.xlsx("Dados Daphnopsis Nov2024.xlsx",colNames=T,detectDates=T,sheet="Maio2022")

year2023<-read.xlsx("Dados Daphnopsis Nov2024.xlsx",colNames=T,detectDates=T,sheet="Maio2023")

year2024<-read.xlsx("Dados Daphnopsis Nov2024.xlsx",colNames=T,detectDates=T,sheet="Junho2024")


```

# Data check

-   Colunas importantes

    -   Tag_Num -\> Old_tag (2022 -\> 2023)
    -   New_Tag (A partir de 2023)
    -   Plot
    -   Site
    -   Sexo
    -   Est_Rep (Estg_Vida em 2022)
    -   DAS_x e DAS_y
    -   Height

-   Checking columns

    -   Columns are not necessarily equal between years, which reflects the changes decision on data collection.
        -   What data is collected has changed but not how data was collected!

-   Checking Tags

    -   A new tag replaced tags used first
    -   Also check if there is no repeated tags. "-" and "Nova" means no old tags (i.e. tags of first census)

```{r label="Data checking", echo=FALSE, eval=F}

#file.edit("Scripts/Data checking.R")

```

# Data standardization

-   Sum branches when necessary
    -   Two branches are written as (3+1), so, total DAS is 4.
-   Mean DAS
-   Transform year2022\$Tag_Num to character

```{r label="Data cleaning"}
source("Scripts/Data cleaning.R")
```

# Merge data from different years

# Statistics

-   Individuals pear year
    -   2022: `r length(year2022$Tag_Num)`
    -   2023 `r length(year2022$Tag_Num)`
    -   2024 `r length(year2022$Tag_Num)`

```{r label="Number of individuals per year"}
data.frame(Years=c("2022","2023","2024"),
      Indiv=c(length(year2022$DAS),
        length(year2023$DAS),
        length(year2024$DAS)))
```



# Old scripts

```{r "Script para sexagem"}
file.edit("Script para sexagem.r")

```
